---
title: "across BLOGPOST"
author: "Jen Richmond"
date: "24/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have been playing with the new across() function in dplyr and working out how to get nicely formatted summary stats tables. 

# load packages

```{r}
library(tidyverse)
library(palmerpenguins)
library(skimr)
library(gt)
library(tableone)
```

# get data

```{r}
penguins <- penguins

```

# The goal

```{r}
penguin_bits <- c("bill", "flipper")
mean <- c(43.92, 200.92)
sd <- c(5.46, 14.06)
min <- c(32.1, 172)
  max <- c(59.6, 231)

df <- data.frame(penguin_bits, mean, sd, min, max)

df %>%
  gt::gt()
```

I want to get summary stats for bill and flipper length (mean, sd, min, and max) and have them display in a nice table like this. This manual df construction is a bit silly?  Is there an easy way to get summary stats in this format? 


# create summary stats the old way
```{r}
penguins %>%
  summarise(bill_mean = mean(bill_length_mm, na.rm = TRUE),
            bill_sd = sd(bill_length_mm, na.rm = TRUE),
             bill_min = min(bill_length_mm, na.rm = TRUE),
             bill_max = max(bill_length_mm, na.rm = TRUE), 
             flipper_mean = mean(flipper_length_mm, na.rm = TRUE),
            flipper_sd = sd(flipper_length_mm, na.rm = TRUE),
             flipper_min = min(flipper_length_mm, na.rm = TRUE),
             flipper_max = max(flipper_length_mm, na.rm = TRUE))

```
The summarise function spits out summary stats in SUPER wide format. 

# create summary stats using across()

The new across() function gets you the same thing in many fewer lines of code, but still, it is VERY wide. 
```{r}

penguins %>%
  summarise(across(contains("length"), list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE)) 

```

# skimr

skim is nice because it is tidyverse compatible (so you can select with contains) but you can also make the output a dataframe so you can mess with it. 
```{r}
skimtable <- penguins %>%
  select(contains("length")) %>%
  skim()

skimtable
  
```

skim gives you more than you need so there is some selecting and renaming to do after the fact. 
```{r}

names(skimtable)

skimtable_rename <- skimtable %>%
  select(skim_variable, numeric.mean, numeric.sd, numeric.p0, numeric.p100) %>%
  rename(penguin_bits = skim_variable, mean = numeric.mean, sd = numeric.sd, min = numeric.p0, max = numeric.p100)

gt::gt(skimtable_rename)
```


# tableone package

The TableOne package gives you mean and SD, I wonder whether there is a way to add other summary stats (like min & max to the TableOne?)


```{r}


to_summarise <- c("bill_length_mm", "flipper_length_mm")

CreateTableOne(vars = to_summarise, data = penguins)
```



